comgen-outline

### 2017-06-12 paper/comgen-paper.Rmd
habituals prior data and local path in paper
habituals model figures

d.hab.priors <- read.csv("/Users/mht/Documents/research/habituals/data/priors/priors-2.csv")

@@

d.hab.endorsement.bayes <- d.hab.endorsement %>% 
  group_by(habitual, logAnnualRate) %>%
  summarize(k = sum(response), n = n()) %>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         low  = qbeta(.025, a, b),
         high = qbeta(.975, a, b),
         MAP_h = (a-1)/(a+b-2),
         mean = a / (a + b))

fig.habituals.endorsement.vs.freq <- 
  ggplot(d.hab.endorsement.bayes, aes( x = logAnnualRate, y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_jitter(width = 0.1, shape = 21)+
  ylab("Human endorsement probability")+
  xlab("Frequency of target")+
  scale_x_continuous(limits = c(-1,6), 
                     breaks = c(-1, 0, 0.7, 2.5, 4, 5.9),
                     labels = c("three years", "annually", "bi-annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-0.01, 1.01), 
                     breaks = c(0, 0.5, 1))+
  theme(axis.text.x = element_text(angle = 90))
```

@@

d.hab.endorse.regression <- left_join(
  d.hab.endorsement.bayes,
  unique(select(
    d.hab.endorsement, habitual, logAnnualRate, mixture
  )) %>%
    mutate(regression_freq = predict(glm.hab.endorse.freq, ., type = "response"), 
           regression_freq_distinct = predict(glm.hab.endorse.freq.distinct, ., type = "response"))
) %>%
  gather(src, prediction, regression_freq, regression_freq_distinct) %>%
  mutate(sqErr = (MAP_h-prediction)^2)

ggplot(d.hab.endorse.regression,
       aes (x = prediction, y = MAP_h, ymin = low, ymax = high, color = mixture))+
  geom_errorbar(alpha = 0.1)+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  # geom_text_repel(data = d.hab.endorse.regression.freq.distinct %>%
  #                   filter(sqErr > 0.1),
  #                 aes(label = paste(habitual, time_period)), force = 5, size = 3)+
  geom_point()+
  xlim(0,1)+
  ylim(0,1)+
  coord_fixed()+
  xlab("Logistic model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src)

@@

```{r habituals-fullmodel}
n_chains <- 1
n_samples <- 10000
burn <- n_samples / 2
lag <- 5
i <- 2
model_prefix <- "results-habituals-S1-ffPrior2-3components-normalPrior-silenceAlt-"

m.samp <- data.frame()
for (i in seq(1, n_chains)){
  mi <- fread(paste(project.path,  "models/habituals/results/", model_prefix, 
                    n_samples, "_burn", burn, "_lag", lag, "_chain", i, ".csv", sep = ""))
  m.samp.i <- mi %>% mutate(chain = i) %>% rename(habitual = item, time_period = param)
  m.samp <- bind_rows(m.samp, m.samp.i)
  #print(i)
}

m.hab.somemodel.endorsement <- fread(
  paste(project.path,"models/habituals/results/", 
        "results-habituals-fixedThreshold-S1-ffPrior2HyperStructured-silenceAlt-10000_burn5000_lag5_chain1.csv",
        sep = "")
  ) %>% mutate(chain = 1) %>% rename(habitual = item, time_period = param) %>%
  filter(type == 'predictive') %>%
  group_by(type, habitual, time_period, cat, gender) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

m.hab.fullmodel.endorsement <- m.samp %>%
  filter(type == 'predictive') %>%
  group_by(type, habitual, time_period, cat, gender) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

m.hab.endorse.rsa <- bind_rows(
  left_join(
    d.hab.endorsement.bayes,
    m.hab.fullmodel.endorsement
  ) %>%
    mutate(src = "generics_model"),
  left_join(
    d.hab.endorsement.bayes,
    m.hab.somemodel.endorsement
  ) %>%
    mutate(src = "some_model")
)

habituals.endorsement.models <- bind_rows(
  d.hab.endorse.regression %>% 
  mutate(prediction_lower = prediction,
         prediction_upper = prediction),
  left_join(
    m.hab.endorse.rsa %>% 
    rename(prediction = MAP, prediction_lower = cred_lower, prediction_upper = cred_upper),
    d.hab.endorse.regression %>% 
    select(habitual, time_period, log_times, distinctiveness)
  )
) %>%
  mutate(src = factor(src, levels = c( 
                                      "generics_model",
                                      "some_model",
                                      "regression_freq", 
                                      "regression_freq_distinct"
                                     ),
                      labels = c(
                                 "Speaker model - uncertain semantics",
                                 "Speaker model - fixed semantics",
                                 "Frequency alone", "Frequency + Distinctiveness"
                                 ))) %>%
  filter(src == "Speaker model - uncertain semantics") %>%
  ggplot(., aes ( x = prediction, xmin = prediction_lower, xmax = prediction_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = log_times ))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(shape = 21)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src, nrow = 2)+
  theme(legend.position = "bottom")

# with(m.hab.endorse.rsa %>% 
#   filter(src == "generics_model"), cor(MAP_h, MAP, use='pairwise.complete.obs'))^2

habituals.endorsement.models
```

>>>>

d.hab.priors <- read.csv(paste(project.path, "data/habituals/priors/", "habituals-priors.csv", sep = ""))

@@

d.hab.endorsement.bayes <- d.hab.endorsement %>% 
  group_by(habitual, logAnnualRate, time_period) %>%
  summarize(k = sum(response), n = n()) %>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         low  = qbeta(.025, a, b),
         high = qbeta(.975, a, b),
         MAP_h = (a-1)/(a+b-2),
         mean = a / (a + b))

example.habituals <- c("climbs mountains_year", "hikes_year", "runs_year",
                       "goes to the movies_month", "smokes cigarettes_month")
fig.habituals.endorsement.vs.freq <- 
  ggplot(d.hab.endorsement.bayes, aes( x = logAnnualRate, y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_jitter(width = 0.1, shape = 21)+
  ylab("Human endorsement probability")+
  xlab("Frequency of target")+
  geom_label_repel(data = d.hab.endorsement.bayes %>% 
                    mutate(sentence = paste(habitual, time_period, sep = "_")) %>%
                    filter(sentence %in% example.habituals), aes( label = habitual ),
                       fontface = 'bold', color = 'white',
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.color = 'grey50')+
  scale_x_continuous(limits = c(-1,6), 
                     breaks = c(-1, 0, 0.7, 2.5, 4, 5.9),
                     labels = c("three years", "annually", "bi-annually", "monthly", "weekly", "daily"))+
  scale_y_continuous(limits = c(-0.01, 1.01), 
                     breaks = c(0, 0.5, 1))+
  theme(axis.text.x = element_text(angle = 90))

@@

d.hab.endorse.regression <- left_join(
  d.hab.endorsement.bayes,
  unique(select(
    d.hab.endorsement, habitual, logAnnualRate, time_period, mixture
  )) %>%
    mutate(regression_freq = predict(glm.hab.endorse.freq, ., type = "response"), 
           regression_freq_distinct = predict(glm.hab.endorse.freq.distinct, ., type = "response"))
) %>%
  gather(src, prediction, regression_freq, regression_freq_distinct) %>%
  mutate(sqErr = (MAP_h-prediction)^2)

# ggplot(d.hab.endorse.regression,
#        aes (x = prediction, y = MAP_h, ymin = low, ymax = high, color = mixture))+
#   geom_errorbar(alpha = 0.1)+
#   geom_abline(intercept = 0, slope = 1, lty = 3)+
#   # geom_text_repel(data = d.hab.endorse.regression.freq.distinct %>%
#   #                   filter(sqErr > 0.1),
#   #                 aes(label = paste(habitual, time_period)), force = 5, size = 3)+
#   geom_point()+
#   xlim(0,1)+
#   ylim(0,1)+
#   coord_fixed()+
#   xlab("Logistic model prediction")+
#   ylab("Human habitual endorsement")+
#   facet_wrap(~src)

@@

```{r habituals-fullmodel}
n_chains <- 1
n_samples <- 10000
burn <- n_samples
lg <- 5
i <- 1
model_prefix <- "results-habituals-jointModel-S1-"

m.hab.samp <- data.frame()
for (i in seq(1, n_chains)){
  mi <- fread(paste(project.path,  "models/habituals/results/", 
                    model_prefix, "smtncs_habitual-",
                    n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
  m.samp.i <- mi %>% mutate(chain = i)
  m.hab.samp <- bind_rows(m.hab.samp, m.samp.i)
}

n_chains <- 1
n_samples <- 1000
burn <- n_samples
lg <- 5
i <- 1
m.hab.fixed.samp <- data.frame()
for (i in seq(1, n_chains)){
  mi <- fread(paste(project.path,  "models/habituals/results/", 
                    model_prefix, "smtncs_some-",
                    n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
  m.samp.i <- mi %>% mutate(chain = i)
  m.hab.fixed.samp <- bind_rows(m.hab.fixed.samp, m.samp.i)
}

m.hab.somemodel.endorsement <- m.hab.fixed.samp %>%
  filter(type == 'predictive') %>%
  rename(habitual = B, time_period = D, binned_freq = E) %>%
  group_by(habitual, time_period, binned_freq) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

m.hab.fullmodel.endorsement <- m.hab.samp %>%
  filter(type == 'predictive') %>%
  rename(habitual = B, time_period = D, binned_freq = E) %>%
  group_by(habitual, time_period, binned_freq) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

m.hab.endorse.rsa <- bind_rows(
  left_join(
    d.hab.endorsement.bayes,
    m.hab.fullmodel.endorsement
  ) %>%
    mutate(src = "habituals_model"),
  left_join(
    d.hab.endorsement.bayes,
    m.hab.somemodel.endorsement
  ) %>%
    mutate(src = "some_model")
)

habituals.endorsement.models <- bind_rows(
  d.hab.endorse.regression %>% 
  mutate(prediction_lower = prediction,
         prediction_upper = prediction),
  left_join(
    m.hab.endorse.rsa %>% 
    rename(prediction = MAP, prediction_lower = cred_lower, prediction_upper = cred_upper),
    d.hab.endorse.regression %>% 
    select(habitual, time_period, logAnnualRate, mixture)
  )
) %>%
  mutate(src = factor(src, levels = c( 
                                      "habituals_model",
                                      "some_model",
                                      "regression_freq", 
                                      "regression_freq_distinct"
                                     ),
                      labels = c(
                                 "Speaker model - uncertain semantics",
                                 "Speaker model - fixed semantics",
                                 "Frequency alone", "Frequency + Distinctiveness"
                                 ))) %>%
  ggplot(., aes ( x = prediction, xmin = prediction_lower, xmax = prediction_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.7)+
  geom_errorbarh(alpha = 0.7)+
  geom_point(shape = 21)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human habitual endorsement")+
  facet_wrap(~src, nrow = 2)+
  theme(legend.position = "bottom")

```
```{r habituals-endorsement-figure, fig.width=11, fig.height=8}
grid.arrange(fig.habituals.endorsement.vs.freq, 
             habituals.endorsement.models,
             generics.endorsement.insets, ncol = 2,
             layout_matrix = cbind(c(1,1,3), c(2,2,3)))
```
